---
title: "Get started with reaborn"
description: >
  Install reaborn, paste seaborn code into R, and extend the result with ggplot2.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with reaborn}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  dev = "ragg_png", dpi = 96, fig.width = 6.5, fig.height = 4.6, fig.retina = 2,
  message = FALSE, warning = FALSE
)
library(reaborn)
set.seed(0)
```

**reaborn** is an R port of Python's [seaborn](https://seaborn.pydata.org),
built on [ggplot2](https://ggplot2.tidyverse.org). It mirrors seaborn's function
API exactly and renders visually indistinguishable plots — and because every
result is a real `ggplot`, you can keep extending it with the grammar of graphics.

## Install

```r
# install.packages("remotes")
remotes::install_github("shawntz/reaborn")
```

## One import sets the scene

Attaching reaborn does three things, mirroring `import seaborn as sns` followed
by `sns.set_theme()`:

```{r}
library(reaborn)
```

1. It **sets the seaborn theme and palette globally**, so every plot inherits the
   familiar seaborn look.
2. It exposes **`sns.`-prefixed aliases** for every function, so pasted Python runs verbatim.
3. It binds the Python literals **`True`**, **`False`**, and **`None`** to R's
   `TRUE`, `FALSE`, and `NULL`.

## Your first plot

Load one of seaborn's example datasets and make a plot. This is literally seaborn
syntax — string column names, named arguments, the `sns.` prefix:

```{r}
penguins <- load_dataset("penguins")

sns.scatterplot(data = penguins, x = "bill_length_mm", y = "bill_depth_mm",
                hue = "species")
```

Prefer idiomatic R? Drop the `sns.` prefix — the bare names work too:

```{r}
histplot(data = penguins, x = "flipper_length_mm", hue = "species",
         multiple = "stack", kde = TRUE)
```

## Every plot is a ggplot

This is reaborn's superpower over seaborn. A plotting call returns a `ggplot`
object, so you can layer on facets, scales, themes, and extra geoms:

```{r}
scatterplot(data = penguins, x = "bill_length_mm", y = "bill_depth_mm", hue = "species") +
  ggplot2::facet_wrap(~island) +
  ggplot2::scale_x_log10() +
  ggplot2::labs(title = "Penguin bills, by island")
```

## The function families

reaborn implements all ~40 seaborn functions. A few entry points:

| Goal | Function(s) |
|---|---|
| Relationships between numeric variables | `scatterplot()`, `lineplot()`, `relplot()` |
| Distributions | `histplot()`, `kdeplot()`, `ecdfplot()`, `displot()` |
| Categorical comparisons | `boxplot()`, `violinplot()`, `barplot()`, `stripplot()`, `catplot()` |
| Model fits | `regplot()`, `lmplot()` |
| Matrices | `heatmap()`, `clustermap()` |
| Multi-plot grids | `pairplot()`, `jointplot()` |

See the [Gallery](gallery.html) for live examples of each, and the
[function reference](https://shawntz.github.io/reaborn/reference/) for full
argument lists.

## Coming from seaborn?

In most cases you change **nothing but the language host**. After
`library(reaborn)`, the `sns.` aliases, the global theme, and the
`True`/`False`/`None` literals are all in scope.

| Python (seaborn) | R (reaborn) |
|---|---|
| `import seaborn as sns` | `library(reaborn)` |
| `sns.set_theme()` | *automatic on load* |
| `sns.scatterplot(data=df, x="a", y="b", hue="g")` | same line, verbatim |
| `True` / `False` / `None` | `True` / `False` / `None` (bound to `TRUE`/`FALSE`/`NULL`) |
| `[1, 2, 3]` · `{"a": 1}` · `(1, 2)` | `c(1, 2, 3)` · `list(a = 1)` · `c(1, 2)` |

The one thing that's truly different — and better — is what you do *after* the
call: instead of mutating a matplotlib `Axes`, you add ggplot2 layers with `+`.
